---
title: "Homework 1" 
author: "Will Scheib"
fontsize: 12pt
geometry: margin=1in
urlcolor: black
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, highlight=FALSE)
```

## Problem 1  

### a
```{r}
a <- seq(4751, 4835, 7)
a
```

### b
```{r}
b <- rep("Night", 5)
b
```

### c
```{r}
c <- seq(89, 101, 1)
c
```

### d
```{r}
d <- rep(seq(21, 18, -1), rep(3, 4))
d
```

### e
```{r}
e <- seq(75, 57, -3)
e
```

### f
```{r}
f <- c(seq(25, 75, 10), seq(70, 50, -5))
f
```

### g
```{r}
g <- rep(seq(400, 500, 25), seq(5, 1, -1))
g
```

### h
```{r}
h <- rep(seq(3, 1, -1), 5)
h
```

### i
```{r}
i <- c(seq(124, 68, -8), seq(63, 38, -5))
i
```
### j
```{r}
j <- rep(c("Morning", "Afternoon", "Night"), 2)
j
```  

## Problem 2 

### a
```{r}
pnorm(26.24, mean = 63, sd = 11)
```
The probability that a runner runs this 10k faster than 26.24 minutes is approximately 0.0004%. 

### b
```{r}
qnorm(0.05, mean = 63, sd = 11)
```
The runner should have a goal time of about 42 minutes and 30 seconds.

### c
```{r}
29000*(1-pnorm(90, mean = 63, sd = 11))
```
I would expect about 205 runners to have their races disrupted.

## Problem 3 

### a
```{r}
1-pbinom(0, size = 12, prob = 0.2)
```
The probability that the lie detector indicates that at least one of the applicants is lying is 93.1%. 

### b
```{r}
input <- 0:11
results <- round(1-pbinom(input, size = 12, prob = 0.2), 3)
results
```

### c
```{r}
desc <- 1:12
desc_results <- rbind(desc, results)
desc_results
```
The descriptor values also show 3 decimal places. 

### d
```{r}
names(results) <- desc
results
```

### e
The labeled vector is easier to read because the label numbers stay in a sensible format. 

### f
```{r}
results_df <- data.frame(desc=desc, prob=results)
results_df
```

### g
```{r}
results_df2 <- data.frame(prob=results)
results_df2
```

### h
The second one, because the data.frame already has row labels, which are essentially what the descriptors vector contains.  

## Problem 4   
```{r}
name <- "Gretchen Martinet"
dep <- "Statistics"
courses <- c("STAT 2559", "STAT 3080")
names(courses) <- c("Course1", "Course2")
act <- c(FALSE, TRUE)
names(act) <- c("Course1", "Course2")
enr <- list(Course1 = 5, Course2 = c(75, 90, 90))
days <- list(
  Course1 = c("Tuesday", "Thursday"),
  Course2 = cbind(c("Monday", "Tuesday", "Tuesday"),
                  c("Wednesday", "Thursday", "Thursday"))
)

mylist <- list(Name = name, Department = dep, Courses = courses,
               ActiveTeach = act, Enr = enr, Days = days)
mylist
```