---
title: "Homework 8" 
author: "Will Scheib"
fontsize: 12pt
geometry: margin=1in
urlcolor: black
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, highlight=FALSE)
library(magrittr)
library(tidyverse)
```

```{r}
set.seed(06072000)
```

## Problem 1
### a
```{r}
data1 <- read.csv("data/data1.csv")
pop.cor.data1 <- cor(data1$V1, data1$V2)
pop.cor.data1
```

### b
```{r}
ttest.paired <- function(data) {
  samp <- data[sample(nrow(data), 13),]
  t.test(samp[,1], samp[,2], mu=0, alternative="two.sided", paired=TRUE)$p.value < 0.05
}
paired.data1 <- mean(replicate(10000, ttest.paired(data1)))
paired.data1
```

### c
```{r}
ttest.two.samp <- function(data) {
  samp <- data[sample(nrow(data), 13),]
  t.test(samp[,1], samp[,2], mu=0, alternative="two.sided")$p.value < 0.05
}
two.samp.data1 <- mean(replicate(10000, ttest.two.samp(data1)))
two.samp.data1
```

\newpage
## Problem 2
### a
```{r}
data2 <- read.csv("data/data2.csv")
pop.cor.data2 <- cor(data2$V1, data2$V2)
pop.cor.data2
```

### b
```{r}
paired.data2 <- mean(replicate(10000, ttest.paired(data2)))
paired.data2
```

### c
```{r}
two.samp.data2 <- mean(replicate(10000, ttest.two.samp(data2)))
two.samp.data2
```

\newpage
## Problem 3
### a
```{r}
data3 <- read.csv("data/data3.csv")
pop.cor.data3 <- cor(data3$V1, data3$V2)
pop.cor.data3
```

### b
```{r}
paired.data3 <- mean(replicate(10000, ttest.paired(data3)))
paired.data3
```

### c
```{r}
two.samp.data3 <- mean(replicate(10000, ttest.two.samp(data3)))
two.samp.data3
```

\newpage
## Problem 4
```{r}
results.table <- cbind(
  data1=c(pop.cor=pop.cor.data1, paired=paired.data1, two.samp=two.samp.data1),
  data2=c(pop.cor=pop.cor.data2, paired=paired.data2, two.samp=two.samp.data2),
  data3=c(pop.cor=pop.cor.data3, paired=paired.data3, two.samp=two.samp.data3)
)
results.table
```
For two data sets that are symmetric and positively correlated, the Type I error for a two sample t-test will be smaller than 0.05. When they are negatively correlated, the Type I error for a two sample t-test will be larger than 0.05. When they are not correlated at all, the Type I error for a two sample t-test will be around 0.05. In all cases, the Type I error for a paired t-test is around 0.05.

\newpage
## Problem 5
### a
```{r}
data4 <- read.csv("data/data4.csv")
pop.cor.data4 <- cor(data4$V1, data4$V2)
pop.cor.data4
```

### b
```{r}
paired.data4 <- mean(replicate(10000, ttest.paired(data4)))
paired.data4
```

### c
```{r}
two.samp.data4 <- mean(replicate(10000, ttest.two.samp(data4)))
two.samp.data4
```

\newpage
## Problem 6
### a
```{r}
data5 <- read.csv("data/data5.csv")
pop.cor.data5 <- cor(data5$V1, data5$V2)
pop.cor.data5
```

### b
```{r}
paired.data5 <- mean(replicate(10000, ttest.paired(data5)))
paired.data5
```

### c
```{r}
two.samp.data5 <- mean(replicate(10000, ttest.two.samp(data5)))
two.samp.data5
```

\newpage
## Problem 7
### a
```{r}
data6 <- read.csv("data/data6.csv")
pop.cor.data6 <- cor(data6$V1, data6$V2)
pop.cor.data6
```

### b
```{r}
paired.data6 <- mean(replicate(10000, ttest.paired(data6)))
paired.data6
```

### c
```{r}
two.samp.data6 <- mean(replicate(10000, ttest.two.samp(data6)))
two.samp.data6
```

\newpage
## Problem 8
```{r}
results.table <- cbind(
  data4=c(pop.cor=pop.cor.data4, paired=paired.data4, two.samp=two.samp.data4),
  data5=c(pop.cor=pop.cor.data5, paired=paired.data5, two.samp=two.samp.data5),
  data6=c(pop.cor=pop.cor.data6, paired=paired.data6, two.samp=two.samp.data6)
)
results.table
```
For two data sets that are skewed and positively correlated or not correlated, the Type I error for a two sample t-test will be smaller than 0.05. When they are negatively correlated, the Type I error for a two sample t-test will be larger than 0.05.

For two data sets that are skewed and positively correlated, the Type I error for a paired t-test will be around 0.05. When they are negatively correlated, the Type I error for a paired t-test will be larger than 0.05. When they are not correlated at all, the Type I error for a paired t-test will be smaller than 0.05.
